Belowground sink limitation alters growth and carbon balance of *Eucalyptus* seedlings

### 2015-12-14 manuscript/manuscript_pve.Rmd
abbreviation changes (added to notes).
manuscripts edits made to all documents and in thesis word document

---
title: |
  | Chapter 2 
  | Belowground sink limitation alters growth and carbon balance of Eucalyptus seedlings
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

<!-- \ -->
<!-- Courtney E. Campany^1^, Belinda Medlyn^1^, Remko A. Duursma^1^. -->

<!-- \ -->
<!-- ^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia -->

<!-- \ -->
<!-- *Corresponding author*:  -->
<!-- Courtney Campany -->
<!-- E: courtneycampany@gmail.com -->

@@

```{r, echo=FALSE}
allomfig <- "2.1"
airvarsfig <- "2.2"
allocfig <- "2.3"
asatfig <- "2.4"
achemfig <- "2.5"
massmodelfig <- "2.6"
massmodelfig2 <- "2.S1"

pvetab1 <- "2.1"
pvetab2 <- "2.2"
pvetab3 <- "2.3"
pvetabS <- "2.S3"

@@

#2.1  Abstract

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A~n~*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A~n~* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A~n~*, these reductions were not large enough to explain observed growth responses. Thus, as *A~n~* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

<!-- ## Keywords {.unnumbered} -->
<!-- photosynthesis, growth, sink regulation, carbon allocation, soil volume -->

#2.2  Introduction

Understanding plant growth and its relationship to C assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales leaf photosynthesis (*A~n~*) and respiratory losses together determine net C balance and will correlate to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue C storage pools. This has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by *A~n~*, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of C supply and growth of woody species.

\
In woody species, the coordination of *A~n~* and growth has been studied with manipulations of C source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A~n~* [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A~n~* of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A~n~* in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A~n~* do not always infer similar responses in growth.

\
Alternatively, manipulating plant tissue C sinks is often used to investigate the correlation of *A~n~* and growth. This is because metabolic signaling networks, relaying information on C and N status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of *A~n~* occurs, a close coordination between declines in *A* and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling and low temperatures at high elevations. In these studies, down regulation of *A* was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in *A~n~* were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, *A~n~*, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A~n~* is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower C exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and *A~n~* are coordinated.

\
This study utilizes a novel field design to investigate the coordination between growth and *A~n~* in *Eucalyptus tereticornis* Sm. seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to isolate the effect of restricted soil volume. We used freely-rooted seedlings as a control for the container size treatments. Empirical results were combined with a simple plant growth model to simulate seedling growth with a C mass balance approach, which was then compared to observed harvested seedling mass. The model used whole-plant C gain, scaled from instantaneous rates of leaf *A~n~*, to quantify seedling dry mass production over the 120 day experiment.

@@

4). Last, observed seedling mass was expected to correspond to growth model mass predicted from a simple C balance model taking into account measured rates of photosynthesis.

#2.3  Materials and Methods

##2.3.1 Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils at this site are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

@@

##2.3.2 Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

##2.3.3 Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (A~sat~) and saturating light and [CO~2~] (A~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 05^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). A~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and A~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol photons m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (g~s~) were logged 5 times and averaged for both A~sat~ and A~max~.

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15~th~, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

\
Leaf dark respiration rates (R) was measured on each seedling during the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of R are standardized rates at 25 &deg;C using a Q~10~ value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

##2.3.4 Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

##2.3.5 Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of N content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of N in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the VPDP scale.

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf N content. Similarly, TNC-free leaf N (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC and leaf N content.

##2.3.6 Seedling growth model
We developed a simple seedling growth model that utilized leaf *A* rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

\
(1)
$$P_i = L\left(\frac{C_{day'i}~\sigma_s}{\epsilon_c}\right) - R$$
where L is total plant leaf area (m^2^), C~day,i~ is the predicted daily carbon assimilation (g C g mass^-1^ d^-1^), $\sigma$~s~ is a self shading parameter (explained below), $\epsilon$~c~ is a biomass conversion efficiency parameter (g C g mass^-1^) and R is the mass based total respiration of all tissue components (g C d^-1^). Total respiration was calculated as

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is published or local tissue respiration rates of fine roots, coarse roots or stems (Table `r pvetabS`, g C g mass^-1^ d^-1^) and M~c~ is the standing biomass of each component (g). R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_c}{dt} = A_{c}~P_{i}~-~(\Lambda_{c}~M_{c})$$
where A~c~ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$~stem~ was assumed to equal 0.  Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

>>>>

---
title: Belowground sink limitation alters growth and carbon balance of *Eucalyptus* seedlings
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes

@@

\
Courtney E. Campany^1^, Belinda Medlyn^1^, Remko A. Duursma^1^.

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
*Corresponding author*:
Courtney Campany
E: courtneycampany@gmail.com

@@

```{r, echo=FALSE}
allomfig <- "1"
airvarsfig <- "2"
allocfig <- "3"
asatfig <- "4"
achemfig <- "5"
massmodelfig <- "6"
massmodelfig2 <- "S1"

pvetab1 <- "1"
pvetab2 <- "2"
pvetab3 <- "3"
pvetabS <- "S3"

@@

#Abstract

Interpreting limitations to plant growth requires understanding of the balance between carbon (*C*) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, *C* allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A*~n~) to allocate daily *C* uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A*~n~ explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A*~n~, these reductions were not large enough to explain observed growth responses. Thus, as *A*~n~ and growth were not tightly coordinated, excess photosynthetic *C* not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf *C* uptake.

## Keywords {.unnumbered}
photosynthesis, growth, sink regulation, carbon allocation, soil volume

#Introduction

Understanding plant growth and its relationship to *C* assimilation requires knowledge of the mass balance that must be achieved between *C* uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net *C* balance and will correlate to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue *C* storage pools. This has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by *A*~n~, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of *C* supply and growth of woody species.

\
In woody species, the coordination of *A*~n~ and growth has been studied with manipulations of *C* source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A*~n~ [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A*~n~ of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], *C* limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant *C* gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A*~n~ do not always infer similar responses in growth.

\
Alternatively, manipulating plant tissue *C* sinks is often used to investigate the correlation of *A*~n~ and growth. This is because metabolic signaling networks, relaying information on *C* and *N* status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of *A*~n~ occurs, a close coordination between declines in *A*~n~ and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling and low temperatures at high elevations. In these studies, down regulation of *A*~n~ was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground *C* sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, *A*~n~, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease *C* sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A*~n~ is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower *C* exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between *C* storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and *A*~n~ are coordinated.

\
This study utilizes a novel field design to investigate the coordination between growth and *A*~n~ in *Eucalyptus tereticornis* Sm. seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to isolate the effect of restricted soil volume. We used freely-rooted seedlings as a control for the container size treatments. Empirical results were combined with a simple plant growth model to simulate seedling growth with a *C* mass balance approach, which was then compared to observed harvested seedling mass. The model used whole-plant *C* gain, scaled from instantaneous rates of leaf *A*~n~, to quantify seedling dry mass production over the 120 day experiment.

@@

4). Last, observed seedling mass was expected to correspond to growth model mass predicted from a simple *C* balance model taking into account measured rates of photosynthesis.

#Materials and Methods

##Experimental design
This experiment was located at the Hawkesbury Forest Experiment site in Richmond, NSW, Australia. Plots were located in an open cover paddock that was converted from native pasture grasses. Top soils at this site are an alluvial formation of low-fertility sandy loam soils (380 and 108 mg kg^-1^ total N and phosphorus respectively) with low organic matter (0.7 %) and low water holding capacity. At this site a soil hard layer exists at ~1.0 m with a transition to heavy clay soils. The climate for the region is classified as sub-humid temperate.

@@

##Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

##Photosynthetic parameters
Leaf gas exchange measurements were performed fortnightly at saturating light (*A*~sat~) and saturating light and [CO~2~] (*A*~max~) on new fully expanded leaves. Measurements were initiated only after sufficient new leaf growth occurred (March 05^th^, 2013), approximately 6 weeks following planting, and continued until the biomass harvest. Leaf level gas exchange was measured with a standard leaf chamber (2 x 3 cm) equipped with blue-red light emitting diodes using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA). *A*~sat~ measurements were made at PPFD of 1800 $\mu$mol m^-1^ s^-1^ and [CO~2~] of 400 $\mu$l l^-1^ and *A*~max~ with [CO~2~] of 1600 $\mu$l l^-1^ and PPFD of 1800 $\mu$mol photons m^-1^ s^-1^. This choice of light level to achieve light saturation is consistent with other studies on *Eucalyptus* species [@kallarackal1997ecophysiological;@pinkard1998photosynthetic;@crous2013photosynthesis;@drake2014capacity]. These measurements were conducted during midday (10:00-14:00 h) with leaf temperature maintained at 25 &deg;C. After CO~2~ and water vapor flux values stabilized in the leaf chamber, net CO~2~ assimilation rate and stomatal conductance (*g*~s~) were logged 5 times and averaged for both *A*~sat~ and *A*~max~.

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15~th~, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, *J*~max~ and *Vc*~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

\
Leaf dark respiration rates (*R*) was measured on each seedling during the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

##Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and *R*. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

##Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of *N* content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of *N*. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of *N* in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf *N* content. Similarly, TNC-free leaf *N* (*N*~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC and leaf *N* content.

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*P*~i~) is then given by

\
(1)
$$P_i = L\left(\frac{C_\mathrm{day'i}~\sigma_\mathrm{s}}{\epsilon_\mathrm{c}}\right) - R$$
where *L* is total plant leaf area (m^2^), *C*~day,i~ is the predicted daily carbon assimilation (g C g mass^-1^ d^-1^), $\sigma$~s~ is a self shading parameter (explained below), $\epsilon$~c~ is a biomass conversion efficiency parameter (g C g mass^-1^) and *R* is the mass based total respiration of all tissue components (g C d^-1^). Total respiration was calculated as

\
(2)
$$R = \Sigma(R_\mathrm{c}~M_\mathrm{c})$$
where *R*~c~ is published or local tissue respiration rates of fine roots, coarse roots or stems (Table `r pvetabS`, g C g mass^-1^ d^-1^) and *M*~c~ is the standing biomass of each component (g). *R*~leaf~ is represented in the calculation of *C*~day~ (described below). The change in individual component biomass (*M*~c~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_\mathrm{c}}{dt} = A_\mathrm{c}~P_\mathrm{i}~-~(\Lambda_\mathrm{c}~M_\mathrm{c})$$
where *A*~c~ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$~stem~ was assumed to equal 0. Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

\
*C*~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package in R with the mean photosynthetic parameters (*J*~max~, *Vc*~max~, R and *g*~1~) for each treatment and meteorological data from an onsite weather station. *J*~max~ and *Vc*~max~ were estimated from AC~i~ curves (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] with observed *g*~s~ values. Methods of the coupled leaf gas exchange model are described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *C*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *C*~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *C*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.
